A Mountaintop View Requires Minimal Sorting: \break A Faster Contour Tree Algorithm \footnote{The full updated version of this paper is available online \url{\si{http://web.engr.illinois.edu/\string~\si{raichel2}/\si{paint.pdf}}}}

### 2014-11-04 paint-contour-tree.tex
valid changed

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%
%%%%%%%%%%Comment out the switch below for full version 
%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\def \STOC {}

@@

Our algorithm requires numerous ideas: partitioning the contour tree into join and split trees, 
a local growing procedure to iteratively build contour trees, and the use of heavy path 
decompositions for the time complexity analysis. There is a crucial use of a family of binomial heaps to 
maintain priorities, ensuring that any comparison made is between nodes in the same path in the contour tree.
We also prove lower bounds showing that the $\sum_{p\in P} |p|\log|p|$ complexity
is inherent to computing contour trees. 
%Consider the family of instances where our algorithm
%requires $C$ comparisons. Any algorithm computing contour trees on this family requires
%$\Omega(C)$ comparisons in the worst case.

@@

We complement \Thm{main-alg} with a lower bound, suggesting that the comparison complexity
of $\sum_{p \in P(T)} |p|\log |p|$ is inherent for any contour tree algorithm. The exact phrasing
of the lower bound requires some care. Consider the $d=2$ (terrain) case.
Let a path decomposition $P$ be called \emph{valid} if it is actually output by the algorithm
on some input.
%As referenced in \Thm{main-alg}, for any $f:\MM \mapsto \RR$ our deterministic algorithm defines a 
%specific path decomposition $P(T)$%
\ignore{\FullVer{
\footnote{The specific decomposition used is defined in \Sec{pathDecomp}, 
but also depends on the raining and painting procedures described in \Sec{rain} and \Sec{paint}.  
The important thing is that our algorithm is deterministic, and so for a fixed $\MM$ and $f$ it is well defined.}.}
\SubmitVer{
\footnote{The specific decomposition used is defined in the full version of the paper.  
The important thing is that our algorithm is deterministic, and so for a fixed $\MM$ and $f$ it is well defined.}.}
We define $\cost(f) = \sum_{p \in P(T)} |p|\log |p|$ for this specific decomposition $P(T)$.
Ideally, we would like to argue that any contour-tree algorithm
must pay $\cost(f)$ for ant $f$. 
But this is obviously false, since one could consider an algorithm
that hard codes the answer for $f$ and merely checks this answer.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%
%%%%%%%%%%Comment out the switch below for full version 
%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\def \STOC {}

@@

Our algorithm requires numerous ideas: partitioning the contour tree into join and split trees, 
a local growing procedure to iteratively build contour trees, and the use of heavy path 
decompositions for the time complexity analysis. There is a crucial use of a family of binomial heaps to 
maintain priorities, ensuring that any comparison made is between nodes in the same descending path in the contour tree.
We also prove lower bounds showing that the $\sum_{p\in P} |p|\log|p|$ complexity
is inherent to computing contour trees. 
%Consider the family of instances where our algorithm
%requires $C$ comparisons. Any algorithm computing contour trees on this family requires
%$\Omega(C)$ comparisons in the worst case.

@@

We complement \Thm{main-alg} with a lower bound, suggesting that the comparison complexity
of $\sum_{p \in P(T)} |p|\log |p|$ is inherent for any contour tree algorithm. The exact phrasing
of the lower bound requires some care. Consider the $d=2$ (terrain) case.
%
%
We define a class of valid path decompositions.  This include any decomposition of rooted trees,
and more generally is a large superset of all decomposition produced by our 
algorithm.
%\footnote{Our construction likely works for any decomposition, however, 
%to keep the presentation simple we restrict to a subset relevant for our algorithm.}
%
%
%Let a path decomposition $P$ be called \emph{valid} if it is actually output by the algorithm
%on some input.
%
%
%As referenced in \Thm{main-alg}, for any $f:\MM \mapsto \RR$ our deterministic algorithm defines a 
%specific path decomposition $P(T)$%
\ignore{\FullVer{
\footnote{The specific decomposition used is defined in \Sec{pathDecomp}, 
but also depends on the raining and painting procedures described in \Sec{rain} and \Sec{paint}.  
The important thing is that our algorithm is deterministic, and so for a fixed $\MM$ and $f$ it is well defined.}.}
\SubmitVer{
\footnote{The specific decomposition used is defined in the full version of the paper.  
The important thing is that our algorithm is deterministic, and so for a fixed $\MM$ and $f$ it is well defined.}.}
We define $\cost(f) = \sum_{p \in P(T)} |p|\log |p|$ for this specific decomposition $P(T)$.
Ideally, we would like to argue that any contour-tree algorithm
must pay $\cost(f)$ for ant $f$. 
But this is obviously false, since one could consider an algorithm
that hard codes the answer for $f$ and merely checks this answer.
